---
title: "CV_analysis"
author: "Moaraj_Hasan"
date: "29 Januar 2017"
output:
  html_document: default
  pdf_document: default
---

Generate a Visual Basic Script that can convert xls and xlsx files into csv in local directory 
xls files are notoriously difficult to handle in R, csv conversion allows low memory

```{r setup,}
#setwd("D:/Dropbox/Aging BXD Study/D _ Metabolomics/D _ Protocol Optimization/CV Calcs")

excel_script <- 
    'if WScript.Arguments.Count < 2 Then
    WScript.Echo "Please specify the source and the destination files. Usage: ExcelToCsv <xls/xlsx source file> <csv destination file>"
Wscript.Quit
End If

csv_format = 6

Set objFSO = CreateObject("Scripting.FileSystemObject")

src_file = objFSO.GetAbsolutePathName(Wscript.Arguments.Item(0))
dest_file = objFSO.GetAbsolutePathName(WScript.Arguments.Item(1))

Dim oExcel
Set oExcel = CreateObject("Excel.Application")

Dim oBook
Set oBook = oExcel.Workbooks.Open(src_file)

oBook.SaveAs dest_file, csv_format

oBook.Close False
oExcel.Quit'

script_file_name = "ExcelToCsv.vbs"
write(excel_script,file = script_file_name)
# The script above allows command line conversion of xlxs file to csv conversiton in command line
#The script syntax: 
#XlsToCsv.vbs [sourcexlsFile].xls [destinationcsvfile].csv

```

The following script takes thes first xls file and performs the conversion using the script
```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tools)
abs_path <- file_path_as_absolute(dir(pattern = "\\.xls")[1])
cmd_command <- paste(c(script_file_name, abs_path, 
                       paste(strsplit(abs_path,".xls"),".csv", sep = "")), 
                     sep = " ", collapse = " ")

system(command = cmd_command)
```

Install all the required dependcies
```{r, message=FALSE, warning=FALSE, eval=FALSE}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("plyr")
install.packages("matrixStats")
install.packages("readr")
install.packages("tidyr")
install.packages("splitstackshape")
```

Initalize data handling dependencies
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(plyr)
library(readr)
library(tidyr)
library(splitstackshape)
library(matrixStats)
library(tibble)
library(heplots)
```

Function for Splitting columns based on substring
```{r}
cSplit <- function(indt, splitCols, sep = ",", direction = "wide", 
                   makeEqual = NULL, fixed = TRUE, drop = TRUE, 
                   stripWhite = FALSE) {
  message("`cSplit` is now part of the 'splitstackshape' package (V1.4.0)")
  ## requires data.table >= 1.8.11
  require(data.table)
  if (!is.data.table(indt)) setDT(indt)
  if (is.numeric(splitCols)) splitCols <- names(indt)[splitCols]
  if (any(!vapply(indt[, splitCols, with = FALSE],
                  is.character, logical(1L)))) {
    indt[, eval(splitCols) := lapply(.SD, as.character),
         .SDcols = splitCols]
  }
  
  if (length(sep) == 1) 
    sep <- rep(sep, length(splitCols))
  if (length(sep) != length(splitCols)) {
    stop("Verify you have entered the correct number of sep")
  }

  if (isTRUE(stripWhite)) {
    indt[, eval(splitCols) := mapply(function(x, y) 
      gsub(sprintf("\\s+%s\\s+|\\s+%s|%s\\s+", 
                   x, x, x), x, y), 
      sep, indt[, splitCols, with = FALSE], 
      SIMPLIFY = FALSE)]
  }  
  
  X <- lapply(seq_along(splitCols), function(x) {
    strsplit(indt[[splitCols[x]]], split = sep[x], fixed = fixed)
  })
  
  if (direction == "long") {
    if (is.null(makeEqual)) {
      IV <- function(x,y) if (identical(x,y)) TRUE else FALSE
      makeEqual <- ifelse(Reduce(IV, rapply(X, length, how = "list")),
                          FALSE, TRUE)
    }
  } else if (direction == "wide") {
    if (!is.null(makeEqual)) {
      if (!isTRUE(makeEqual)) {
        message("makeEqual specified as FALSE but set to TRUE")
        makeEqual <- TRUE
      }
      makeEqual <- TRUE
    } else {
      makeEqual <- TRUE
    }
  }
  if (isTRUE(makeEqual)) {
    SetUp <- lapply(seq_along(X), function(y) {
      A <- vapply(X[[y]], length, 1L)
      list(Mat = cbind(rep(seq_along(A), A), sequence(A)),
           Val = unlist(X[[y]]))
    })    
    Ncol <- max(unlist(lapply(SetUp, function(y) y[["Mat"]][, 2]), 
                       use.names = FALSE))
    X <- lapply(seq_along(SetUp), function(y) {
      M <- matrix(NA_character_, nrow = nrow(indt), ncol = Ncol)
      M[SetUp[[y]][["Mat"]]] <- SetUp[[y]][["Val"]]
      M
    })
    if (direction == "wide") {
      X <- lapply(seq_along(X), function(x) {
        colnames(X[[x]]) <- paste(splitCols[x], 
                                  sequence(ncol(X[[x]])), 
                                  sep = "_")
        X[[x]]
      })
      if (isTRUE(drop)) {
        cbind(indt, do.call(cbind, X))[, eval(splitCols) := NULL][]
      } else {
        cbind(indt, do.call(cbind, X))
      }
    } else {
      indt <- indt[rep(sequence(nrow(indt)), each = Ncol)]
      X <- lapply(X, function(y) as.vector(t(y)))
      indt[, eval(splitCols) := lapply(X, unlist, use.names = FALSE)][]
    }  
  } else {
    Rep <- vapply(X[[1]], length, integer(1L))
    indt <- indt[rep(sequence(nrow(indt)), Rep)]
    indt[, eval(splitCols) := lapply(X, unlist, use.names = FALSE)][]
  }
}
```

Import the Annotated files into R and perform the condition groups
```{r, eval=FALSE}

#View(Data_Curated_Annotations)

#Parses Meta Data in File
meta_data_vector <- cSplit(as.data.frame(names(Data_Curated_Annotations))
                           ,"names(Data_Curated_Annotations)", " ")
meta_data_vector$`names(Data_Curated_Annotations)_2` <- NULL
names(meta_data_vector) <- c("index", "extraction", "mouse_id", "diet_id", "sex")

#Makes Labels for MetaData and ions
meta_levels <- c("index", "extraction", "mouse_id", "diet_id", "sex")
ionMz <- as.character(Data_Curated_Annotations$ionMz) 
metab_data_names <- c(meta_levels, ionMz)

#Transpose Data Frame

#removes the first three rows and sets ion mass as column nam
metab_data <- t(Data_Curated_Annotations)
metab_data <- data.frame(meta_data_vector, metab_data)
row.names(metab_data) <- NULL

names(metab_data) <- metab_data_names
metab_data <- metab_data [-c(1,2,3),]

#Covert Data Frame to numeric
metab_data <-lapply(metab_data[6:length(names(metab_data))], function(x) as.numeric(x))
metab_data <- data.frame(meta_data_vector[-c(1:3)], metab_data)

```

Import Annotated Files, Tranposed, all commas, spaces and semi colons removed in notepad++. 
INSERT "index, extraction, mouse, diet, sex," to first row

```{r}
Data_Curated_Annotations_transpose <- read_csv("Data_Curated_Annotations_transpose.csv", 
    col_types = cols( `71.0137` = col_integer(), extraction = col_character(), 
        index = col_integer()))

metab_data <- Data_Curated_Annotations_transpose %>% dplyr::mutate_each_(., funs = funs(as.factor), vars = c("extraction", "diet", "sex", "mouse")) 


```

Get the Mean reading of each metabolite and visualize in a box plot
```{r, message=FALSE, warning=FALSE}

sample_sd <-lapply(metab_data[6:length(names(metab_data))], function(x) sd(x))
sample_mean <-lapply(metab_data[6:length(names(metab_data))], function(x) mean(x))

p <- ggplot(metab_data, aes(factor(extraction), mouse ))
p + geom_violin() + geom_boxplot(width=.1)

  
extraction <- metab_data[metab_data$extraction %in% "H1",]


```

Generate Table to compare Chow and High Fat Diet metabolites
Script for extracting just cohort subsets
```{r}
#Exctraction Cohorts
Hot_data  <- filter(metab_data,extraction == "Hot") %>% filter(sex == "F")
H1_data   <- filter(metab_data,extraction == "H1") %>% filter(sex == "F")
H24_data  <- filter(metab_data,extraction == "H24") %>% filter(sex == "F")
NH24_data <- filter(metab_data,extraction == "NH24") %>% filter(sex == "F")

#Diet Cohorts
HF_data <- filter(metab_data, diet == "HF") %>% filter(sex == "F")
CD_data <- filter(metab_data, diet == "CD") %>% filter(sex == "F")
```

Add indexes for replicaites using the mod function on the sample index
```{r}
#mod<-function(x,m){
#    t1<-floor(x/m)
#    return(x-t1*m)
#}

#index_vector        <- c(1:nrow(metab_data))
#replicate_index     <- mod(metab_data$index,2)
#replicate_index     <- as.data.frame(as.character(rev((replicate_index + 1))))
#index_vector        <- cbind(index_vector, replicate_index)
#names(index_vector) <- c("index", "replicate")

metab_data <- merge(metab_data 
                    %>% mutate(replicate  = paste("rep_", rep(c(1,2), times = dim(metab_data)[1]/2), sep = "")) 
                    %>% select(index, replicate), metab_data, by = "index")

#zou need to have unqiue values in the bz mere
#can be done by selecting multiple columns
```


ANOVA fit
```{r, fig.height=4, fig.width=4}

Data_Annotations    <- read.csv(file = "Data_Curated_Annotations.csv",stringsAsFactors = FALSE)
primary_names       <- Data_Annotations$Primary.Name
meta_data           <- c("index","replicate","extraction","mouse","diet","sex")
primary_meta_names       <- c(meta_data, primary_names)

#names(metab_data)   <- primary_meta_names
#duplicated_colnames <- which(duplicated(primary_meta_names,fromLast = FALSE))
#metab_data<-lapply(duplicated_colnames, function(x) metab_data[-x])


duplicated_cols <- which(duplicated(primary_meta_names))
filtered_metab_data <- metab_data %>% select(-duplicated_cols)
filtered_primary_meta_names <- primary_meta_names[-duplicated_cols]
final_metab_data <- filtered_metab_data %>% setnames(filtered_primary_meta_names)

#Set up variables for the aov analysis
metab <- final_metab_data

counts <- log(as.matrix(metab[6:10]))
#extraction  <- metab$extraction
#mouse   <- metab$mouse
#diet    <- metab$diet
#sex     <- metab$sex
#replicate <- metab$replicate
  
fit <- aov(formula = as.formula("counts ~ sex + diet + extraction + replicate"), data = metab)
fit2 <- glm(formula = as.formula("counts ~ sex + diet + extraction + replicate"), data = metab)

metab.mod <- lm(as.matrix(metab[7:10]) ~ sex + diet, data=metab)
Anova(metab.mod, test.statistic="Roy")


summary(fit)
summary(fit2)

plot(fit,which = 1:6)
plot(fit2, which = 1:6)

p <- ggplot(metab_data, aes(x = diet, y = metab_data$Pyruvaldehyde)) +
  geom_boxplot(fill = "grey80", colour = "blue") +
  scale_x_discrete() + xlab("Treatment Group") +
  ylab("Log Metabolite Intensities")

```


Violin Plots for the Metabolite Cohorts
```{r}
HF_mean <-lapply(HF_data[6:length(names(HF_data))], function(x) mean(x))
CD_mean <-lapply(CD_data[6:length(names(CD_data))], function(x) mean(x))

HF_sd <-lapply(HF_data[6:length(names(HF_data))], function(x) sd(x))
CD_sd <-lapply(HF_data[6:length(names(HF_data))], function(x) sd(x))

Diet_data <- data.frame(as.numeric(HF_mean), as.numeric(CD_mean))
names(Diet_data) <- c("HF_mean","CD_mean")
Diet_data$Index <- 1:length(CD_mean)

metab_data <- melt(Diet_data, id="Index")

q <- ggplot(metab_data, aes(factor(variable), log(value)))
q + geom_violin() + geom_boxplot(width=.1)
```

compare the technical replicates
```{r}


#x <- metab_data[metab_data$mouse == 1037,]
#y <- [x$extraction == "H1",]
#z <- x[x$extraction == "H1",]

qqplot(as.numeric(z[1,6:length(z)]),as.numeric(z[2,6:length(z)]))
abs_diff <- colDiffs(as.matrix(z[,6:length(z)]))
t.test(as.numeric(z[1,6:ncol(z)]),as.numeric(z[2,6:ncol(z)]))


mouse_levels <- levels(metab_data$mouse)
extraction_levels <- levels(metab_data$extraction)
diet_levels <- levels(metab_data$diet)

stat_df <- data.frame(
     first_condition = character(),
     second_condition = character(),
     t_test = numeric(),
     stringsAsFactors = FALSE
                 )

for(i in 1:length(mouse_levels))   {
     x <- mouse_levels[i]
     print(i)
     
     for (j in 1:length(extraction_levels)){
          y <- extraction_levels[j]
          z <- filter(metab_data, mouse==x) %>% filter(extraction == y)
          
          if (nrow(z)==2)          {
               xx <- c(mouse_levels[i],(extraction_levels[j]))
               t_test_result  <- t.test(as.numeric(z[1,6:ncol(z)]),
                                        as.numeric(z[2,6:ncol(z)]))
               
               result_vector <- c(xx,t_test_result$p.value)
               
               abs_diff <- colDiffs(as.matrix(z[,6:length(z)]))
               result_vector <- c(result_vector, abs_diff)

               stat_df <- rbind(stat_df, result_vector)
               
               }
          
          else if (nrow(z) == 0)   {
               xx <- c(as.character(mouse_levels[i]),
                       as.character(extraction_levels[j]))
               length(xx) <- ncol(z)
               stat_df <- rbind(stat_df, xx)
               }
               }
               }


```

Get the Mean reading of each metabolite and visualize in a box plot
```{r}
#Last row has a classifier and the rowMeans operation must be done on numeric columns
metab_data[,"Mean"] <- rowMeans(metab_data[,6:ncol(metab_data)-1])

# Name[ row, column ] - Roq SD Funtions
st_dev <- vector(mode = "numeric", length = (nrow(metab_data)-1))
for(i in 6:nrow(metab_data)) {
  st_dev[i] <- sd(Hot_data[i,1:ncol(Hot_data)])
}

Hot_data[,"ST Dev"] <- st_dev

df <- data.frame(annotation_file$Primary.Name,annotation_file$ionMz, Hot_data, Hot_data$`ST Dev`)
names(df) <- c("Name","ionMZ","Mean", "SD")

df2 <- df %>% arrange(Mean)
```




